Measurement of total hadronic differential cross sections in the LArIAT experiment

### 2018-04-11 Chapter-1/1-Theory.tex
s

From a group theory perspective, the SM has a rather complex group structure, where a gauge group is formed with the direct product of other three groups as shown in eq. \ref{eq:SMGroup}. Drawing a parallel with the electroweak symmetry breaking mechanism, where the $ SU(2)_T \otimes U(1)_Y$ is recovered from $U(1)_{EM}$, an interesting line of simplification for the SM group structure would be to think about a similar mechanism where $SU(3)_C  \otimes SU(2)_T \otimes U(1)_Y$ is recovered from an hypothetical larger group. Just as the electroweak unification becomes evident at energies higher than the Higgs VEV,  a direct manifestation of Grand Unification Theories (GUTs) would occur at even higher energies.

As the smallness of neutrino masses suggests to the existence of a higher mass scale \textcolor{red}{IS THIS TRUE?}, an other, even stronger, hint to Grand Unification comes from the slope of running of the coupling constants. The coupling constants for the electromagnetic, weak and strong interactions in the SM vary as a function of the interaction energy as shown in figure \ref{fig:constants}; they do not exactly meet, but their trend is interesting enough to push theorists to construct theories where perfect unification is achieved with the addition of new particles.

@@

\subsubsection{Nucleon decay}\label{sec:theoryPDK}
Baryon number is accidentally conserved in the Standard Model. Even though no baryon number violation has been experimentally observed thus far, no underlying symmetry in line with the Noether paradigm \cite{Noether1971} explains its conservation. Almost all Grand Unified Theories predict at some level baryon number violation in the form of nucleon decay on long time-scales.  Given the impossibility to reach grand unification energy scales with collider experiments ($\sqrt{s} > 10^{15}$ GeV),  an indirect proof of GUT is needed. The experimental observation of nucleon decay may be the only viable way to explore these theories and it is therefore a subject of great interest \cite{Adams:LBNE}. %Both experiments and theory indeed suggest the energy scale for convergence of the running coupling constants of the Standard Model to be over $10^{15}$ GeV. This energy scale seems impossible to access by any foreseeable accelerator experiment, leaving baryon number violation  to be the only testable process.

>>>>

From a group theory perspective, the SM has a rather complex group structure, where a gauge group is formed with the direct product of other three groups as shown in eq. \ref{eq:SMGroup}. Drawing a parallel with the electroweak symmetry breaking mechanism, where the $ SU(2)_T \otimes U(1)_Y$ is recovered from $U(1)_{EM}$, an interesting line of simplification for the SM group structure would be to devise a similar mechanism where $SU(3)_C  \otimes SU(2)_T \otimes U(1)_Y$ is recovered from an hypothetical larger group. Just as the electroweak unification becomes evident at energies higher than the Higgs VEV,  a direct manifestation of Grand Unification Theories (GUTs) would occur at even higher energies.

As the smallness of neutrino masses suggests the existence of a higher mass scale, an other, even stronger, hint to Grand Unification comes from the slope of running of the coupling constants. The coupling constants for the electromagnetic, weak and strong interactions in the SM vary as a function of the interaction energy as shown in figure \ref{fig:constants};  they cannot exactly meet under the current experimental constraints, but their trend is interesting enough to push theorists to construct theories where perfect unification is achieved with the addition of new particles. Both experiments and theory indeed suggest the energy scale for convergence of the running coupling constants of the Standard Model to be over $10^{15}$ GeV. This energy scale seems impossible to access by any foreseeable accelerator experiment.%, leaving baryon number violation  to be the only testable process.

%Models I want to mention: SUSY and SU(5) kaon modes and p -> epi0

%The $SU(5)$ group is the smallest simple group containing $SU(3)_C  \otimes SU(2)_T \otimes U(1)_Y$. Georgi and Glashow proposed the ``minimal SU(5)"

@@

\subsubsection{Nucleon decay}\label{sec:theoryPDK}
Baryon number is accidentally conserved in the Standard Model. Even though no baryon number violation has been experimentally observed thus far, no underlying symmetry in line with the Noether paradigm \cite{Noether1971} explains its conservation. Almost all Grand Unified Theories predict at some level baryon number violation in the form of nucleon decay on long time-scales.  Given the impossibility to reach grand unification energy scales with collider experiments ($\sqrt{s} > 10^{15}$ GeV),  an indirect proof of GUT is needed. The experimental observation of nucleon decay may be the only viable way to explore these theories and it is therefore a subject of great interest \cite{Adams:LBNE}.
